---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Note that the following 3 variables are treated as outcome variables and will not be considered to be input variables for cluster analysis
  .	"Satisfaction on the insurance firm's customer service"
  .	"Satisfaction on the insurance policy"
  .	"Will consider switch"


For this assignment, you will complete following steps:

.	Examine the data structure
.	Demonstrate the descriptive information of the variables that you are interested in
    
.	Select numeric variables and categorical variables (mixed data) and use PAM() to conduct cluster analysis
    o	Decide the number of clusters
    o	Decide the best cluster solution
    o	Show the differences of key factors' characteristics across clusters 
    o	Check the differences of the 3 outcome variables across the clusters
    o	Discuss your findings and implications to understand employee profiles and satisfaction

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/ca034330/Google Drive/Corey - School/!Spring 2019 B/Assignment1/data/"))
#knitr::opts_knit$set(root.dir = normalizePath("D:/Google Drive/Corey - School/!Spring 2019 B/Assignment1/data/"))
```

```{r}
library(cluster) 
library(NbClust) 
library(factoextra)
library(dplyr)
library(ggplot2)
library(readxl)
library(Rtsne)
```

```{r data_load}
w <- read_xlsx("Employee profile and satisfaction data.xlsx",col_names = TRUE)
```

```{r}
names(w)
summary(w)
str(w)
```

```{r}
w.numeric<- as.data.frame(w[,c(5,11,13:22,24)])
w.fact<- as.data.frame(w[,c(3,4,7:10,12,23,25)])
w.fact[] <- lapply(w.fact, factor)
```

```{r}
w.fact$'Job Involvement' <- factor(w.fact$'Job Involvement',levels = c("Low", "Medium", "High", "Very High"))
w.fact$'Environment Satisfaction' <- factor(w.fact$'Environment Satisfaction',levels = c("Low", "Medium", "High", "Very High"))
w.fact$'Work Life Balance' <- factor(w.fact$'Work Life Balance', levels = c("Bad", "Good", "Better", "Best"))
w.fact$'Education' <- factor(w.fact$'Education', levels = c("Below College", "College", "Bachelor", "Master", "Doctor"))
```

```{r}
w1<- cbind(w.numeric, w.fact)
```

```{r}
numComplete <- NbClust(w1, distance="euclidean", min.nc=2, max.nc=10,
                       method="complete", index="all")
```


```{r}
gower_dist <- daisy(w1, metric = "gower", type = list(logratio = 3))
```

```{r}
gower_mat <- as.matrix(gower_dist)
```

```{r}
w1[which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]), arr.ind = TRUE)[1, ], ]
```

```{r}
w1[which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]), arr.ind = TRUE)[1, ], ]
```



```{r}
sil_width <- c(NA)
for(i in 2:10){  
  pam_fit <- pam(gower_dist, diss = TRUE, k = i)  
  sil_width[i] <- pam_fit$silinfo$avg.width  
}

plot(1:10, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width)
```

```{r}
k <- 2
pam_fit <- pam(gower_dist, diss = TRUE, k)
pam_results <- w1 %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary
```

```{r}
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)

tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))

ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))
```


```{r}
w.clust <-cbind(w, pam_fit$clustering)
```

```{r}
Freqs <- table(w.clust$"Will consider switch", w.clust$"pam_fit$clustering")
Freqs
```


```{r}
Freqs <- table(w.clust$"Satisfaction on the insurance policy", w.clust$"pam_fit$clustering")
Freqs
```

```{r}
Freqs <- table(w.clust$"Satisfaction on the insurance firm customer service", w.clust$"pam_fit$clustering")
Freqs
```
  .	"Satisfaction on the insurance firm's customer service"
  .	"Satisfaction on the insurance policy"

```{r}
pam(w1, 9, metric = c("euclidean"))
```

```{r}
library(factoextra)
set.seed(1234)
km.res <- eclust(w.clust, "pam", k = 9, graph = FALSE)

table(km.res$cluster)
fviz_cluster(km.res, geom = "point", ellipse.type = "norm",
             palette = "jco", ggtheme = theme_minimal())
```

```{r}
fviz_nbclust(w.clust,FUNcluster = cluster::pam, method = "wss") +
geom_vline(xintercept = 3, linetype = 2)
```


km.res <- eclust(w2, "kmeans", k = 2, nstart = 25, graph = FALSE)
























```{r}
w2<- scale(w1)
summary(w2)

```

```{r}
numComplete <- NbClust(w2, distance="euclidean", min.nc=2, max.nc=10,
                       method="complete", index="all")

```

```{r}
names(numComplete)
```

```{r}
numComplete$Best.nc
```

```{r}
dis = dist(w2, method="euclidean")
```

```{r}
hc = hclust(dis, method="complete")
plot(hc, hang=-1,labels=FALSE, main="Complete-Linkage")
```

```{r}
comp4 <- cutree(hc, 4)
```

```{r}
library(NbClust) 
NbClust(w2, distance="euclidean", min.nc=2, max.nc=10,
        method="ward.D2", index="all")

```

```{r}
hcWard <- hclust(dis, method="ward.D2")

plot(hcWard, labels=FALSE, main="Ward's-Linkage")

ward3 <- cutree(hcWard, 3)

table(ward3)
```

```{r}
table(comp4, ward3)
```

```{r}
aggregate(w2,list(comp4),mean)

par(mfrow=c(1,2))
w3<- w2

```

```{r}
w4<-as.data.frame(cbind(w1, comp4, ward3))

table(w4$comp4,w4$ward3)

```

```{r}
dis = dist(w2, method="euclidean")
hc = hclust(dis, method="complete")

```

```{r}
res.coph <- cophenetic(hc)
```

```{r}
cor(dis, res.coph)
```

```{r}
NbClust(w2, min.nc=2, max.nc=15, method="kmeans")
```

```{r}
set.seed(1234)
km<- kmeans(w2,2,nstart=25)

```

```{r}
table(km$cluster)
```

```{r}
w3<-as.data.frame(cbind(w1, km$cluster))
colnames(w3)[14]<-"clusters"

```

```{r}
aggregate(w3[1:13], by=list(w3$clusters), FUN=mean)
```

```{r}
library(factoextra)
set.seed(1234)
km.res <- eclust(w2, "kmeans", k = 2, nstart = 25, graph = FALSE)

table(km.res$cluster)
fviz_cluster(km.res, geom = "point", ellipse.type = "norm",
             palette = "jco", ggtheme = theme_minimal())

```

```{r}
library(clValid)
```

```{r}
clmethods <- c("hierarchical","kmeans","pam")
intern <- clValid(w2, nClust = 2:4,
                  clMethods = clmethods, validation = "internal" )
```

```{r}
summary(intern)
```

SECTION 1 END HERE!
